On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-04-05 TTau_Halpha/revision01/response_to_ref.txt

Dear Referee,

Thank you very much for reviewing our paper, and for doing so in
timely mannar.  We have reviewed our paper by considering your
comments and suggestions very carefully.  
The answers to your questions follow "*" marks before 
the indented texts below.

@@

===================================================================
General comments

Introduction, 3rd paragraph: The blue-ward asymmetric emission is
generated in magnetospheric accretion models not only by the
occultation of the flow by the stellar photosphere, which occults
mostly high velocity blue-shifted material, but also by the
occultation of part of the receding flow by the inner disk, which
occults mostly low velocity redshifted emission. Both effects (lack of
low velocity red-shifted emission and lack of high velocity
blue-shifted emission) contribute to generate the blue-ward asymmetry.

   * Thank you for point that out. I replaced the sentence 
     from 
	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the stellar photosphere, .... 
     to

	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the stellar photosphere and by the inner
	    part of the accretion disc, .... "

@@

Section 2.1, 2nd paragraph: A better reference to the accretion schock
characteristics would be Calvet & Gullbring (1998) ApJ 509, 802 and
Gullbring et al. (2000) ApJ 544, 927, instead of Konigl (1991) and
Hartmann et al. (1994).

     * I replaced the references as you suggested.

In this paragraph you obtain about 8% of the stellar surface covered
by hot rings. This is quite a large value compared with the values
obtained by Gullbring & Calvet in the papers cited above, which range
rather from 0.1% to 1% for several CTTSs. But this could be due to the
fact that you are using a large Macc rate (10^-7 Msun/yr) for a
CTTS. Does this affect your results ? Was the hot ring surface
coverage recalculated consistently when you computed models with 10^-8
Msun/yr and 10^-9 Msun/yr ? In that case, what values were used ?

   * The surface "area" covered by the hot rings depends only on 
     the geometry of the magnetosphere (Rmi and Rmo); however, 
     the total accretion luminosity (L_acc) depends on the
     mass-accretion rate (Macc).
     The surface intensity of the hot rings is computed according to 
     this total luminosity due to accretion and by assuming they
     radiate as a blackbody.

@@

     If a smaller hot ring surface was adopted, the ring temperature
     will increase in order to have a fixed L_acc for a given amount
     of mass-accretion rate.  So, yes, we calculate the amount of
     total "flux" according to a mass-accretion rate, but the surface
     coverage does not change.

     The level populations are relatively insensitive to the actual
     distribution of the hot spots on the stellar surface and hence,
     the H-alpha line strength is not affected by the change in the
     area coverage of the rings provided that the change is not too
     greate. If the area coverage is much bigger (e.g. 30%), then the
     geometry of the magnetosphere itself must change significantly.
     This would result in the line strength slightly for a fixed
     inclination angle, but overall shape of the profile would not
     change so much. For example, we used Rmi=2.2R* and Rmo = 3.0R*
     (Small and Wide (SW) model of Muzerolle et al 2001) which results
     in 8% hot ring coverage. If we were to use Rmi =2.8R* and
     Rmo=3.0R* (Small and narrow (SN) model of Muzerolle et al 2001),
     then the coverage would be 1% of the total surface, which is more
     similar to the fraction estimated in Gullbring et al (2000).  If
     we look at the model comparison of Muzerolle et al (2001) (see
     their Fig. 5), the line strength changes, but the line profiles
     shape do not change much (except for the 10^3 K model). We
     confirmed, by running additional models, the line shapes do not
     change significantly for H-alpha by changing from SW to SN model.

@@

        "The luminosity ratio and the temperature depend on a adopted
         mass-accretion rate."

Sect. 2.2, 3rd paragraph: You say that the results of Martin (1996)
qualitatively agree with those of Hartmann et al (1994). There is a
comment about that in the Muzerolle et al (1998) (ApJ 492, 743) paper
on page 751, where they also use the Hartmann et al (1994) law. They
say that the Martin (1996) temperature distribution is inadequate to
obtain line emission consistent with the observations, even at Halpha,
while their temperature distribution generates profiles that look like
the observed ones. So I think that you should not say that both
temperatures agree, even qualitatively.  It gives the idea that they
generate the same types of profiles, which is not true. The fact that
they do not agree is actually a big problem, since the Hartmann et
al. (1994) temperature structure, which is somewhat arbitrary,
reproduces the observed profiles, while the Martin (1996) temperature
structure, which was calculated consistently, does not. We are missing
some physics in there.

   * The two temperature structures _do_ agree in a qualitative sense,
     in that their behaviour as a function of distance along a
     streamline follows a similar dependence. Nonetheless, you are
     right that Muzerolle could not reconcile the Martin T structure
     with observations. We have made the following change:

        "Muzerolle et al (1998) found that the line profile models
         computed according to the temperature structure of Martin (1996)
         do not agree with observations, unlike profiles based on the
         (less self-consistent) Hartmann et al. (1994) temperature
         distribution.  It is clear that The temperature structure of
         the magnetosphere is still a large source of uncertainty in
         the accretion model; here we adopted the model of Hartmann et
         al. (1994). "

@@

     http://www.astro.ex.ac.uk/people/rk/profiles/extra2

     To clarify the points you raised above, I have inserted the
     following in text (mostly the same as mentioned above).

     After the 4th paragraph in Section 4.3, the following new
     paragraphs are inserted:

       "Blandford & Payne (1982) showed that in order to launch a wind
        from a disc plane magneto-centrifugally, the launching angle
        (theta_L, measured from the equatorial disc plane) must be
        less than 60 degrees. In our models, the wind launching zone
        is restricted in between Rwi < wi < Rwo (see Fig. 1 and Table
        1) with Rwi = 3 Rstar and Rwo=1 au.  This sets the upper limit
        of the source displacement distance (d) to be around 5.2 Rstar
        if we want to satisfy the launching angle condition of
        Blandford & Payne (1982) for all the stream lines in our disc
        wind model.  In the models presented in Fig. 6 (and later in
        Fig. 9 in next section), we have adopted d =22 Rstar
        instead. We have used this larger source
	displacement distance for the following two reasons.

      	First, we want to simulate a more 'colliminated' wind (toward
	the symmetry axis).  Since the stream lines in our simple disc
	wind are represented by straight lines, the wind never bends
	toward the symmetry axis (z-axis) as the distance from the
	disc increases unlike the model of Blandford & Payne
	(1982). We therefore used a larger launching angle to simulate
	the stream lines in the far field (at a larger distance from
	the disc plane) in their model.  This is an intrinsic problem
	with approximating the stream lines with straight lines.

	Second, the larger source displacement model would provide a
        greater consistency with the fraction of H-alpha profiles
        affected by wind absorption seen in observations.  According
        to the observations of Reipurth (1996, A&AS, 120, 229) and
        Alencar & Basri (2000, ApJ 119, 1881), more than 50 per cent
        of CTTS show wind absorption features in their H-alpha
        profiles. In our model, if the launching angle is too small
        (<60 deg.), the wind absorption feature would not be seen or
        would be very small for profiles computed with i<60 deg., as
        demonstrated in Fig. 7.  The situation will however change if
        the field/stream line is not straight and bending towards the
        symmetry axis as the wind travel higher above the disc.

	In addition to the models given in Figs. 6 and 9, we have 
	computed the models with d=5Rstar (not shown here) which 
	satisfies the wind launching condition of Blandford & Payne 
	(1982) in all parts of the disc wind zone in our model. 
	These models showed that the wind absorption and emission 
	features weaken, but overall morphology dependence of the 
	profiles on the wind temperature  and the wind 
	acceleration rate are similar to the models shown in Figs. 6 
	and 9.  The minimum inclination angle at which the wind 
	absorption becomes important is larger than that of the 
	models with a larger source displacement (d = 22 Rstar) as 
	expected."

@@

     for the profile computed at i=10,55,80 and for beta=0.2, 0.5,
     1.0, 2.0 Twind=7000, 8000, 9000, 10000K

      According to these models which use a more typical
      mass-accretion rate, Type III-B can be seen for at least in
      three models with fast accelerating and high wind temperature
      i.e. (beta, Twind) = (0.2, 10000K), (0.5, 9000K) and (0.5,
      10000K).

      I added a following comments based on these additional runs.

      "Although Type III-B profiles do not occur at macc=10^-8, we
      have examined the intermediate case of macc=3x10-8, which may
      represent a typical T Tauri mass accretion rate  (Gullbring et
      al. 1998; Calvet et al. 2004). We find that the models with fast accelerating,
      high temperature winds do display  Type III-B profiles. It is
      clear that our model, as it stands, requires above-average (but
      perhaps not atypical) mass-accretion rates in order to produce
      Type III-B profiles. It is less clear whether the 14 Type-IIIB
      objects found by Reipurth et al. constitute the highest
      mass-accretors in their sample: the average EW for T~Tauri stars
      with Type III-B profiles is 69\AA, which is in the top quartile
      of their complete T~Tauri star sample, perhaps indicating that
      Type-IIIB profiles are associated with high EWs, and therefore
      strong mass-accretion."

@@

Sect. 5.2, 4th paragraph: The sentece below must be rewritten: ...This
tendency (...) holds for the models with this wind models for wide
ranges of beta...

   * Changed the sentence 
     From 
     "This tendency of the inclination dependency holds for the models
      with this wind models for wide ranges of beta and
      T_wind providing that the emission from the wind is
      greater than that of the magnetosphere."

     To
     "This inclination dependency of EW is seen in the models with
      wide ranges of beta and T_wind, provided that the wind emission
      of the model is dominating the emission from magnetosphere."

**** Additional Corrections *********************************

@@

In addition to the referee's suggestions, we noticed the following 
minor mistakes in the original manuscript, and corrected.

1. The mass-loss rate in figure a1 and figure a2 were miss-labeled in
   the captions.  ==>  corrected.

2. In Figure 13, the velocity ranges of the two plots should be
   same for a better comparison. We changed the range of the 
   right plot to -800km/s <V < 800 km/s.
